Canary in the Coal Mine: Anti-NGO Legislation as a Predictor of Civil Society Repression

### 2021-03-23 analysis/03_models.Rmd
A bunch of changes I made back in March 2020 but then covid happened and I have no idea where I was going with it all so I'll just commit it here and hope for the best

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)

canary <- read_rds(here("data", "derived_data", "canary_clean.rds"))
```

@@

>>>>

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
library(here)
library(brms)
library(lme4)

CHAINS <- 4
ITER <- 2000
WARMUP <- 1000
BAYES_SEED <- 1234
options(mc.cores = parallel::detectCores())  # Use all cores

canary <- read_rds(here("data", "derived_data", "canary_clean.rds"))
canary_lagged <- read_rds(here("data", "derived_data", "canary_clean_lagged.rds"))
```

TODO: (from the envelope on my desk):

1. Null model: PTS ~ (1 | year) + (1 | country)
2. PTS ~ repression + (1 | year) + (1 | country)
3. PTS + repression within/between + (1 | year) + (1 | country)
4. Fuller models with more controls - region dummies vs. (1 | region / country)
5. Switch to ordered logit
6. Logs and SNMM combination
7. Make sure selection on observables is okay
8. IPW?

@@

- Use four versions of each coefficient - within/between year/country
- When interpreting, focus on within (moving up/down from average level). Within country = in a typical country, the effect/association of x on y (i.e. increasing 1 more above average in a normal country is associated with β change in Y). Within year = in a typical year, the effect association of x on y (i.e. increasing 1 more above average in a given year is associated with a β change in Y)

- Within country = increasing 1 unit of X above the average in a typical country is associated with β change in Y
- Within year = increasing 1 unit of X above the average during a typical year is associated with a β change in Y
- Between country = countries with higher average X are associated with a β higher level of Y 
- Between country = years with higher average X are associated with a β higher level of Y

> I'd just go with the basic one unit change marginal effects language. You can talk about it in ranges too since the groups are actually meaningful. The effect of X on Y was twice as great in 2012 as it was in 2008 or something.
https://twitter.com/DavidPoe223/status/1236864935537278976

> When you add a 1 inside (1 + year|continent/country), you are essentially allowing the countries to have their intercepts specified as "global intercept + country-specific displacement from global intecept + continent-specific displacement from global intercept".
https://twitter.com/IsabellaGhement/status/1236849235275829248

> Yes, that model would allow temporal trends specific to each country, while also estimating how trends differ across countries and also across continents relative to the trend for a "typical" country in a "typical" continent (all else being equal).
https://twitter.com/IsabellaGhement/status/1236850446016188416

# Within/between effects (aka Mundlak device)
PTS ~ 1 + CS_repress_demeaned + CS_repress_mean + controls_demeaned + controls_meaned +
  year +  # Allow for year global trend
  (1 + year | country)  # Random country intercepts

https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#model-specification

https://rpubs.com/jebyrnes/causal_mods

http://methods.johndavidpoe.com/2019/02/05/my-talk-at-iu/

```{r}
canary_demeaned <- canary_lagged %>% 
  mutate(gwcode = as.factor(gwcode)) %>% 
  group_by(gwcode) %>% 
  mutate_at(vars(v2csreprss, v2x_polyarchy, gdpcap_log),
            list(between = ~mean(., na.rm = TRUE),  # Between, for countries
                 within = ~. - mean(., na.rm = TRUE)))  # Within, for countries

model_asdf <- lmer(PTS_lead1 ~  # Overall terror/repression level
                     v2csreprss_within + v2csreprss_between +  # Civil society repression
                     v2x_polyarchy_within + v2x_polyarchy_between +  # Democracy
                     gdpcap_log_within + gdpcap_log_between +  # GDP per capita
                     armed_conflict +  # Government involved in conflict with 25+ battle deaths
                     splines::bs(year) +  # Non-linear year trend
                     (1 + v2csreprss_within | region/gwcode) +  # Random slopes for within CS repression
                     (1 + v2csreprss_between | region/gwcode),  # Random slopes for between CS repression
                   data = canary_demeaned)

tidy(model_asdf, effects = "fixed")
coef(model_asdf)$`gwcode:region` %>% rownames_to_column() %>% as_tibble() %>% head()
coef(model_asdf)$region %>% rownames_to_column() %>% as_tibble()
```

```{r}
model_lag1 <- lm(PTS_lead1 ~ barriers_total + v2x_polyarchy + region + gdpcap_log + as.factor(year),
                 data = canary_lagged)
tidy(model_lag1)

modelx_lag1 <- lm(PTS_lead1 ~ v2csreprss + v2x_polyarchy + region + gdpcap_log + as.factor(year),
                 data = canary_lagged)
tidy(modelx_lag1)

modelx_lag1 <- lm(PTS_lead1 ~ v2csreprss,
                 data = canary_lagged)
tidy(modelx_lag1)
```

```{r}
thing1 <- brm(
  bf(PTS_lead1 ~ v2csreprss + v2x_polyarchy + region + gdpcap_log + 
       year + (1 + year | gwcode)),
  family = gaussian, inits = 0,
  control = list(adapt_delta = 0.99),
  data = canary_lagged %>% select(PTS_lead1, v2csreprss, v2x_polyarchy, 
                                  region, gdpcap_log, year, gwcode),
  prior = c(set_prior("normal(0, 10)", class = "Intercept"),
            set_prior("normal(0, 2.5)", class = "b"),
            set_prior("normal(0, 2.5)", class = "sd")),
  chains = CHAINS, iter = 8000, warmup = WARMUP, seed = BAYES_SEED,
  file = here("analysis", "cache", "thing_8000")
)

tidy(thing1)
tidy(thing)

# Random intercepts & random slopes?
lmer(outcome ~ explanatory + controls + (1 | year) + (1 | countrycode))

# Partial pooling?
lmer(outcome ~ explanatory + controls + (1 + year | countrycode))

brm(bf(outcome ~ explanatory + controls + year + (1 + year | countrycode)))

PTS ~ CS_repress + controls + as.factor(year) + 
  (1 + as.factor(year) | countrycode) + (1 + region | )

# Super rich model with country, year, and region effects - information flows through the whole structure: https://twitter.com/statwonk/status/1236794927683698688
PTS ~ CS_repress + controls +
  (1 | continent/country/year)

# Effect of each year and also each country has its own year effects:
(1 | year) + (1 | country) + (1 | year:country)

# Slopes vs. shocks
# https://twitter.com/statwonk/status/1236785673853681664
as.factor(year) + (1 | year) + (1 | country) + (1 | year:country)

PTS ~ CS_repress + controls +
  year + (1 + year | country) +  # Random year slopes within countries
  (1 | continent/country/year)  # Random intercepts nested inside continent/country/year

PTS ~ CS_repress + controls +
  year + (1 + year | country) +  # Random year slopes within countries
  (1 | continent/country/year)  # Random intercepts nested inside continent/country/year

# Nested continent/country effects
PTS ~ CS_repress + controls +
  (1 | year) +  # Allow for year shocks
  (1 + year | continent / country)  # Random nested continent/country intercepts

# Within/between effects (aka Mundlak device)
PTS ~ 1 + CS_repress_demeaned + CS_repress_mean + controls_demeaned + controls_meaned +
  year +  # Allow for year global trend
  (1 + year | continent / country)  # Random nested continent/country intercepts

PTS ~ CS_repress + controls +
  year + (1 | year) +  # YOLO trend + shocks
  (1 + year | continent / country)  # Random nested continent/country intercepts

# Just country effects
PTS ~ CS_repress + controls +
  (1 | year) +  # Allow for year shocks
  (1 + year | country)  # Random nested country intercepts

PTS ~ CS_repress + controls +
  year +  # Allow for year global trend
  (1 + year | country)  # Random nested country intercepts

PTS ~ CS_repress + controls +
  year + (1 | year) +  # YOLO trend + shocks
  (1 + year | country)  # Random nested country intercepts

get_prior(bf(PTS_lead1 ~ v2csreprss + v2x_polyarchy + region + gdpcap_log + (1 | year) + (1 | gwcode)),
  family = gaussian,
  data = canary_lagged)

diff_cs_pts <- brm(
  bf(v2csreprss ~ 0 + PTS_factor, sigma ~ 0 + PTS_factor),
  family = student, 
  data = filter(canary, year == 2010) %>% select(v2csreprss, PTS_factor),
  prior = c(
    # Set group mean prior
    set_prior("normal(0, 1)", class = "b", lb = -4, ub = 4),
    # Set group variance priors. We keep the less informative cauchy(0, 1).
    set_prior("cauchy(0, 1)", class = "b", dpar = "sigma"),
    set_prior("exponential(1.0/29)", class = "nu")
  ),
  chains = CHAINS, iter = ITER, warmup = WARMUP, seed = BAYES_SEED,
  file = here("analysis", "cache", "diff_cs_pts")
)

tidy(thing)

lag1 <- lm(I(sqrt(terrorinclead) - coef(out)["totspendrevlog"]*totspendrevlog) ~ l(totspendrevlog, cow)  + l(govleft, cow) + l(democ, cow) + l(poplog, cow) + l(govcap, cow) + l(conflict, cow) + l(tradelog, cow) + europe +africa + asia + america + sqrt(l(terrorinc, cow))+ sqrt(l(terrorinc, cow, 2))+ as.factor(year), data=reasonabledata)

lag2 <- lm(I(sqrt(terrorinclead) - coef(out)["totspendrevlog"]*totspendrevlog - coef(lag1)["l(totspendrevlog, cow)"]*l(totspendrevlog, cow))~l(totspendrevlog, cow, 2)  + l(govleft, cow, 2) + l(democ, cow, 2) + l(democ, cow, 3) + l(poplog, cow, 2) + l(govcap, cow, 2) + l(conflict, cow, 2) + l(tradelog, cow, 2) + europe +africa + asia + america + sqrt(l(terrorinc, cow, 2))+ sqrt(l(terrorinc, cow,3))+ as.factor(year), data=reasonabledata)
```

### 2021-03-23 analysis/index.Rmd
A bunch of changes I made back in March 2020 but then covid happened and I have no idea where I was going with it all so I'll just commit it here and hope for the best

Many countries across the globe are increasingly cracking down on civil society groups through legal restrictions. While research on democratic backsliding has tended to focus on rights and institutions in the electoral arena, this new wave of laws—many of which appear mundane on their face—represent a bureaucratic form of repressive that could be indicative of more severe human rights abuses and backsliding. This is especially the case for electoral democracies, which unlike autocracies do not aggressively and openly attack civic space. In this paper, we ask: do legal crackdowns on NGOs predict broader civil society repression? We argue that this is the case because anti-NGO laws are among the most subtle means of repression and attract the least least domestic and international condemnation, as many audiences mistake this repression as regulation and are not personally affected by it. These laws also make it easier for states to prevent and deter future large-scale popular mobilization. Using original data on civil society restrictions over the last three decades, we test whether NGO crackdown is a predictor of attacks on journalists, independent media, academic freedom, and freedoms of expression and association. This project has important implications: if crackdown on NGOs is indeed an early warning, it can potentially prevent further deteriorations in space for civil society.

>>>>

Many countries across the globe are increasingly cracking down on civil society groups through legal restrictions. While research on democratic backsliding has tended to focus on rights and institutions in the electoral arena, this new wave of laws—many of which appear mundane on their face—represent a bureaucratic form of repression that could indicate more severe human rights abuses and backsliding. This is especially the case for electoral democracies, which unlike autocracies do not aggressively and openly attack civic space. Do legal crackdowns on NGOs predict broader civil society repression? We argue they do, given that anti-NGO laws are among the most subtle means of repression and attract the least domestic and international condemnation, as many audiences mistake this repression as regulation and are not personally affected by it. Using original data on civil society restrictions over the last three decades, we test whether NGO crackdown is a predictor of attacks on journalists, independent media, academic freedom, and freedoms of expression and association. This project has important implications: if crackdown on NGOs is indeed an early warning, it can potentially prevent further deteriorations in space for civil society.
